---
title: "Assignment 1 - Data Analysis and R Programming"
output: pdf_document
date: "2023-02-10"
author: "Group 5 - Pat, Oat, Shine, Erika, David"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyfoot[CO,CE]{Github Repository - https://github.com/pat-nb/gbc-t405-bus4066-assignment1-r.git}
- \fancyfoot[RE,RO]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 1 - Data Analysis and R Programming

*Author: GBC T405 BUS4066 - Group 5*

*1. Nichapat (Pat) Boonprasertsri, Student ID 101410612*

*2. Chotiros (Oat) Srisiam, Student ID 101411914*

*3. Shine Chen, Student ID 101450231*

*4. Erika Valle, Student ID 101381686*

*5. David Minotas, Student ID 101409821*

***Please import insurance_data.csv before run all code below***

```{r - import csv file or run this code to read csv file}
insurance_data <- read.csv("https://raw.githubusercontent.com/pat-nb/gbc-t405-bus4066-assignment1-r/main/insurance_data.csv")
```

## Data Preprocessing

### Print the structure of dataset

```{r - print the structure of insurance_data}
str(insurance_data)
```

### List the variables in dataset

```{r - list the variables in insurance_data}
names(insurance_data)
```

### Print the top 15 rows of dataset

```{r - print the top 15 rows of insurance_data}
head(insurance_data, n=15)
```

### Write a user defined function using any of the variables from the data set

```{r - create function for calculating year of birth from provided age}
cal_yob <- function(age) {
  2021 - age # assume that 2021 is the year that data is created
}
print(head(cal_yob(insurance_data$age), n = 20))
```

### Use data manipulation techniques and filter rows based on any logical criteria that exist in dataset

```{r - filter patients that age equal to or less than 30 and are diabetic)}
# Attach tidyverse packages to use data manipulation, reading, transforming and visualizing datasets
library("tidyverse")
# Select first 10 rows
insurance_data %>%
  select(age, bmi, diabetic, claim) %>%
  filter(age <= 30, diabetic == "Yes") %>%
  slice_head(n = 10)
```

### Identify the dependent & independent variables and use reshaping techniques and create a new data frame by joining those variables from dataset

```{r - use reshaping techniques and create a new data frame by joining variables}
# Create a new data frame with smoker and diabetic column
reshap_col_smoker_diab <- cbind(insurance_data$smoker, insurance_data$diabetic)
print(head(reshap_col_smoker_diab, n = 10))

# Find and store patients whose age is under 20
df_age_under_20 <- insurance_data %>%
                select(PatientID, age, bmi, diabetic, claim) %>%
                filter(age < 20) %>%
                arrange(age, by_group = TRUE)

# Find and store patients whose age is from 20 to 30
df_age_20_30 <- insurance_data %>%
              select(PatientID, age, bmi, diabetic, claim) %>%
              filter(age >= 20, age < 30) %>%
              arrange(age, by_group = TRUE)

# Create a new data frame with patients whose age under 30 by merging 2 prepared data frame
df_age_under_30 <- rbind(df_age_under_20, df_age_20_30)
print(head(df_age_under_30, n = 10))
```

### Remove missing values in dataset.

```{r - remove records which age are missing}
insurance_data %>% 
  select(2:5, 11) %>%
  filter(!is.na(age)) %>%
  slice_head(n = 10)
```

### Identify and remove duplicated data in dataset

```{r - identify and remove duplicated in insurance_data}
# Identify duplicated data
insurance_data[duplicated(insurance_data)]

# Remove duplicated rows in a data frame
insurance_data %>% 
  select(2:5, 11) %>%
  distinct() %>%
  slice_head(n = 10)

# Remove duplicated rows based on age
insurance_data %>%
  select(2:5, 11) %>%
  distinct(age)
```

### Reorder multiple rows in descending order

```{r - reorder age and claim in descending order}
insurance_data %>%
  select(2:5, 11) %>%
  arrange(-age, -claim) %>%
  slice_head(n = 10)
```

### Rename some of the column names in dataset

```{r - rename PatientID}
insurance_data %>% 
  select(2:5, 11) %>%
  rename(patient_id=PatientID) %>%
  slice_head(n = 10)
```

### Add new variables in data frame by using a mathematical function

```{r - create new variables or column which is year of birth that calculate from age}
insurance_data %>% 
  filter(!is.na(age)) %>%
  mutate(yob = 2021 - age) %>%
  slice_head(n = 10)
```

### Create a training set using random number generator engine

```{r - random select 5% of records}
# Total number of data
count(insurance_data)
# Number of selected 5% data
count(insurance_data) * 0.05

# Random select 5% of records
set.seed(1234)
insurance_data %>%
  sample_frac(0.05, replace = FALSE)
```

### Print the summary statistics of dataset

```{r - find mean claim of each age}
insurance_data %>% 
  group_by(age) %>%
  summarise(mean(claim)) %>%
  slice_head(n = 10)
```

### Use any of the numerical variables from the dataset and perform the following statistical functions: Mean, Median, Mode, Range

```{r - perform statistical functions}
# Find mean of claim
cat("Mean claim = ", as.character(mean(insurance_data$claim)))
# Find median of claim
cat("Median claim = ", median(insurance_data$claim))
# Find mode of age
cal_mode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}
cat("Mode age = ", cal_mode(insurance_data$age))
# Find range of claim
cat("Range claim = ",(range(insurance_data$claim)))
```

## Visualization

### Plot a scatter plot for any 2 variables in dataset

```{r - plot a scatter plot for bmi and claim}
ggplot(data = insurance_data,
       mapping = aes(x=bmi,
                     y=claim)) +
  geom_point()
```

### Plot a bar plot for any 2 variables in dataset

```{r - plot a bar plot for age and smoker factor}
ggplot(data = insurance_data,
       aes(x = age, fill = factor(smoker))) +
  geom_bar()
```

## Correlation

### Find the correlation between any 2 variables by applying least square linear regression model

```{r - find the correlation between claim and bmi}
corr_bmi_claim = cor(insurance_data$claim, insurance_data$bmi, method="pearson")
print(corr_bmi_claim)
```
